Backward Population Synthesis: Mapping the Evolutionary History of Gravitational-Wave Progenitors

### 2023-03-20 src/tex/ms.tex

\begin{abstract}
One promising way to extract information about stellar astrophysics from
gravitational wave catalogs is to compare the catalog to the outputs of stellar
population synthesis modeling with varying physical assumptions.  The parameter
space of physical assumptions in population synthesis is high-dimensional and
the choice of parameters that best represents the evolution of a binary system
may depend in an as-yet-to-be-determined way on the system's properties.  Here
we propose a pipeline to simultaneously infer zero-age main sequence properties
and population synthesis parameter settings controlling modeled binary evolution
from individual gravitational wave observations of merging compact binaries. Our
pipeline can efficiently explore the high-dimensional space of population
synthesis settings and progenitor system properties for each system in a catalog
of gravitational wave observations.  We apply our pipeline to observations in
the third LIGO--Virgo Gravitational-Wave Transient Catalog. We showcase the
effectiveness of this pipeline with a detailed study of the progenitor
properties and population synthesis settings that produce mergers like the
observed GW150914.  Our pipeline permits a measurement of the variation of
population synthesis parameter settings with binary properties, if any;
\textbf{we illustrate the possibility of such capability by presenting
inferences for the recent GWTC-3 transient catalog that suggest that the stable
mass transfer efficiency parameter may vary with primary black hole mass.}  
\end{abstract}

@@

Traditionally, population synthesis studies simulate DCO populations with a
Monte-Carlo approach. Using theoretically motivated distributions of initial
parameters like age, metallicity, component mass, orbital separation, and
eccentricity, initial populations are evolved with fixed astrophysical
assumptions, or population synthesis hyperparameter settings, to produce a
synthetic catalog of DCO mergers which can be compared to parameterized models
derived from observations. \textbf{This method has been applied extensively to
inference on GW data to inform phenomenological models
\citep[e.g.][]{Fishbach2017, Wysocki2019, Callister2021, Delfavero2021,
Farah2022, Delfavero2022}, population synthesis for a single formation
environment \citep[e.g.][]{Belczynski2016, Stevenson2017b, Taylor2018,
Barrett2018, Mastrogiovanni2022} or contributions from multiple formation
environments \citep[e.g][]{Zevin2017, Stevenson2017, Bouffanais2019, Zevin2021,
Wong2021, Bouffanais2021, Mapelli2022}}. These approaches rely on several
assumptions, both in the simulations and model parameterizations fitted to the
GW data, which are unlikely to be fully correct. For example, \textbf{without
knowing whether the parameterization of hyperparameters is sufficiently
accurate}, there is no reason to suppose that the entire population of DCOs
should evolve with the same fixed hyperparameters; there is no a priori reason
the fitted parameterized model should capture the relevant features of the
observed DCO distributions to correspond to the binary physics of interest; etc.
A better approach would be to compare progenitor parameters and hyperparameter
settings directly to DCO observations, treating the population synthesis model
as a mapping from progenitor to merger parameters; merger parameters can then be
mapped into observations using a gravitational waveform family, and parameter
inference proceeds in the usual way, propagating information back up the mapping
chain \citep[e.g.][]{Veitch2015}.  A similar methodology, though with
\emph{fixed} hyperparameter settings, was advanced in
\citet{Andrews2018,Andrews2021}, as we discuss in more detail below.

@@

We find that the progenitor properties and even the ability to produce a
GW150914-like merger event is strongly correlated with hyperparameter settings,
\textbf{consistent with results from prior forward modeling studies cited above};
different assumptions about the black hole masses at merger in GW150914 imply
wildly different formation channels and ZAMS progenitor masses for this event
and some nearby combinations of ZAMS masses are unable, with any reasonable
hyperparameter settings, to produce a DCO merger like GW150914.  We further
exhibit preliminary results of an ongoing analysis over the entire GWTC-3
catalog \citep{GWTC-3} that suggest that the accretion efficiency in stable mass
transfer may depend non-trivially on the primary black hole mass in merging
systems; our methodology allows for a systematic study of such dependencies,
which we will explore more fully in future work.

@@

\subsection{\textbf{Statistical method}}
\label{subsec:stats}
We define the progenitor's zero age main sequence (ZAMS) properties like mass,
orbital period, and eccentricity as progenitor parameters $\bm{\theta'}$, the
parameters that control different uncertain physical prescriptions such as wind
strength and common envelope efficiency as hyperparameters $\bm{\lambda}$, and
random variables that affect certain stochastic process such as whether CO natal
kick unbinds a binary as $\bm{X}$. To avoid clutter in the following derivation,
we denote all the parameters related to mapping a particular progenitor system
into a GW event collectively as evolutionary parameters $\bm{\Theta}$, which
includes $\bm{\theta'}$, $\bm{\lambda}$, and $\bm{X}$.

@@

\begin{align}
    p(\bm{\Theta} | d) = \frac{p(\bm{\theta}(\bm{\Theta})| d) \pi(\bm{\Theta})}{\pi_\mathrm{GW}(\bm{\theta}(\bm{\Theta}))}.
\end{align}
We apply a kernel density estimator with a Gaussian kernel \footnote{We use the
implementation in \texttt{scipy} \cite{2020SciPy-NMeth}} to the posterior
samples released in GWTC-3 to estimate $p(\bm{\theta}(\bm{\Theta})|d)$.
\textbf{A desired prior in the evolutionary parameters can be included through
the prior $\pi(\bm{\Theta})$ in the above equation.}

Including hyperparameters and random variables %in the sampling
drastically increases the dimensionality of the problem, which can make the
sampling process much more computationally expensive to converge. To speed up
the convergence, we first solve a root-finding problem to find points
$\bm{\Theta}$ that will generate GW parameters $\bm{\theta}$ close to the bulk
of the posterior density of a GW event.  We then use those points to initialize
a set of chains in the MCMC process.

@@

In principle, we should only accept solutions that exactly reproduce the LVK
posterior samples in the GW-observable space. However, it is not feasible to
achieve such a condition in practice, therefore we relax the condition in
eq.~\ref{eq:loss} to a small acceptance threshold. For this study, we picked a
threshold of $10^{-2}$. To make sure we find a reasonably complete set of
progenitor parameters that corresponds to the posterior sample point, we use
1000 different initial guesses in solving the optimization problem. As long as
the solution fulfills the acceptance criteria, the root is as valid as all other
roots that fulfill the same acceptance criteria, \emph{regardless to its initial
guess}. This means, we have the freedom to choose the set of initial guesses
however we think might benefit the optimization.

@@

\subsection{\textbf{Binary evolution assumptions}}
\label{subsec:binaries}

@@

\textbf{We follow the critical mass ratio prescription 
described in \cite{Neijssel2019} for donors which fill their Roche lobes during
 other phases of stellar evolution.} 
For systems which begin a common envelope 
with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
stars merge since the donor has likely not formed a strong core-envelope boundary 
\citep{Ivanova2004, Belczynski2008}. \textbf{We fix other assumptions for how stellar and binary 
evolution proceed. In particular, we assume that stars lose mass through stellar winds 
in a metallicity-dependent 
fashion according to \cite{Vink2001,Vink2005}. We assume that binaries are rapidly circularized
during both stable Roche-overflow mass transfer and common envelopes.
We further assume that the CO formation mechanism follows the delayed 
prescription of \citet{Fryer2012} which allows neutrino-driven supernova explosions to 
occur from instabilities which grow on timescales longer than $250\,\rm{ms}$, 
and that natal kicks due to supernovae are reduced by the amount 
of mass which falls back onto the proto-CO
as described in \citet{Fryer2012}.}

\subsection{\textbf{Case-specific configuration}}
\label{subsec:configuration}

We apply the method described in section \textbf{\ref{subsec:stats}} to the first GW event GW150914 \citep{GW150914}.
We use the \texttt{Overall\_posterior} posterior samples for GW150914,
publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}. 
The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
The spin of BHs formed in isolated binaries depends strongly on the angular
momentum transport in massive stars which is still highly uncertain (see however \citet{Fuller2019, Bavera2020}),
therefore we only consider the two component masses in the observable space $\bm{\theta}$ in both the
root-finding stage and the MCMC stage.

For progenitor parameters $\bm{\theta'}$, we characterize each guess with
five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$,
the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not
obtained through the root-finding process because the formation redshift does not intrinsically affect
the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from
formation to the merger and add this time to the lookback time when a GW event is observed and limit the total
time to merger to be less than $13.7\,\rm{Gyr}$.
Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the
Planck 2018 cosmological model \citep{Planck2018}.
The main benefit of post-processing the redshift in this way is that we do not rely on a particular assumption
of star formation rate (SFR) distribution when we solve for the posterior distribution in redshift.
This means we can put the prior on formation redshift (or in general, metallicity-redshfit distribution) in
by reweighting the posterior distribution according to a particular SFR model,
and therefore do not have to rerun the inference every time we change the SFR model. This is especially helpful
since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.

For hyperparameters $\bm{\lambda}$, we choose parameters that strongly affect
binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
the critical mass ratio which determines whether Roche-overflow mass transfer
proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap
$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
employed for CO natal kicks, $\sigma$. 
%Prior

@@

We show a portion of the joint posterior of the progenitor parameters and
hyperparameters of GW150914 in figure \ref{fig:GW150914_posterior}. Each panel
is colored based on the comparison between progenitor parameters (blue),
hyperparameters (orange), or a mix of the two (green). \textbf{The choice
for $\sigma$ does not affect the formation of GW150914-like BBHs.} This is not
unexpected since \textbf{the \citet{Fryer2012} delayed model can significantly reduce} the
natal kick strength at BH formation based on the amount of fallback onto the
proto-CO. In the case of GW150914-like BBHs, the BH progenitors are massive
enough that the fallback reduces the natal kick to zero. Because of this we do
not include $\sigma$ in Figure~\ref{fig:GW150914_posterior}. Similarly, we do
not include the ZAMS orbital period and eccentricity, which are correlated with
one another, but not strongly with the other progenitor parameters or
hyperparameters.

\textbf{The formation redshift distribution is slightly correlated with the $q_{\rm crit, 3}$ and $f_{acc}$ hyperparameters, though the GW150914-like progenitors generally prefer formation redshifts of $z\sim0.1$ (delay times of $\lesssim0.1\,\rm{Gyr}$). This correlation is due to the preference for longer delay times between ZAMS formation and merger for progenitors which undergo stable mass transfer at Roche overflow instead of a common envelope. This result is in direct agreement with \citet{vanSon2022a} who also find that the common envelope evolution channel produces BBH mergers with shorter delay times than the stable Roche-overflow channel.}

\textbf{The joint posterior exhibits} strong correlations between the ZAMS masses, as well as
strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$, and
$f_{\rm{acc}}$ and $\alpha$. We show scatter plots of each of these combinations
in the upper-right inset of the figure.

The correlations between the ZAMS masses are similar to those found by A21 with
the majority of the population preferring primary and secondary masses between
$60$--$90\,M_{\odot}$. However, we also find some ZAMS masses which extend up to
$150\,M_{\odot}$ which is the limit imposed by our assumptions. \textbf{The features in the scatter points of the ZAMS masses, most obvious in the region where $M_1,\rm{ZAMS}\sim 140-150\,M_\odot$, are the outcome of the interplay between the progenitor parameters and hyperparameters. The dominant cause of regions with fewer data points is that fewer binary progenitors produce BBH mergers with chirp masses and mass ratios consistent with GW150914's priors. This is most clearly illustrated for progenitors with primary masses between $142-144\,M_\odot$ and secondary masses below $80\,M_[\odot]$.} We explore the
formation scenarios which lead to successful GW150914-like mergers and those
which fail to produce GW150914-like mergers \textbf{in more detail} below in
Figure~\ref{fig:compare_fixed_variable}.

@@

% Paragraph discussing the potential application on the population side.
To illustrate the potential benefit of using our method on the population level,
we perform the same analysis for all events in GWTC-3. For events announced in
GWTC-1, we use \texttt{Overall\_posterior},
\texttt{PublicationSamples},\texttt{IMRPhenomXPHM\_comoving} and
\texttt{C01:Mixed} posterior samples, respectively. All the data are publicly
available from the Gravitational wave Open Science Center
\citep{LIGOScientific:2019lzm}. Plotted in figure \ref{fig:GWTC-3_f_acc_mass} is
the posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
most\footnote{\textbf{We excluded events for which our reconstructed posterior has KL
divergence greater than 0.1 nats relative to the GWTC-3 posterior; see below.}}
of the events in GWTC-3. Each contour represents the $68\%$ credible interval of
the posterior density for that particular event. There is a suggestive trend
showing that $f_{\rm acc}$ could increase as the mass of the progenitor
increases. Such a trend implies the stable mass transfer phase of a less massive
binary would be preferentially non-conservative, with more conservative mass
transfer in more massive binary systems.

\textbf{We have validated the robustness of every events shown in
\ref{fig:GWTC-3_f_acc_mass} by reprojecting the posterior density in the
evolutionary parameter space to the observable space.} Note that some events did
not pass the KL divergence test we proposed in section \ref{sec:method}; we do
not include these systems in figure \ref{fig:GWTC-3_f_acc_mass}. Binaries that
form low-mass compact objects have lower amounts of fallback and tend to have
correspondingly larger variance in their random natal kicks. Larger kicks can
unbind the progenitor binary, or lead to wide binaries that do not merge within
a Hubble time.  In these cases, the extra variance during the reprojection can
produce a posterior that may not agree with the original posterior, hence
yielding a higher KL divergence.

@@

We present a new pathway to understand binary evolution with GW events in this
paper. Instead of forward modelling an assumed distribution of initial binaries
to the observed population, we find the corresponding evolutionary parameters
event-by-event. In this first work, we showcase the power of the proposed method
with an application to GW150914. We show the joint posterior of the event's
progenitor parameters and hyperparameters.  Our work is a \textbf{data-driven}
way to study GW event progenitors, and also allows the possibility of
constraining the astrophysics related to binary evolution, especially by
capturing the correlation between hyperparameters in different systems.

Our results suggest that the accretion efficiency during stable mass transfer
may depend on the primary black hole mass.  In general, our method returns the
joint posterior distribution of progenitor parameters and hyperparameters for
each event, which enables a data-driven way to study the distribution of
hyperparameters.  That is, once we have a catalog of events, each has their own
posterior in the hyperparameter space, and we can employ well-known techniques
such as hierarchical Bayesian analysis to fit a population model to the
distribution of hyperparameters; this avoids making overly-specific assumptions
like fixing the hyperparameters for all types of event. \textbf{In this work we
focus on illustrating the possibililty of conducting such analysis with our
pipeline. Note that there are concerns that need to be addressed before one can
make statments about the physics, such as the number of channels considered and
specific choices of parameterization. These concerns are beyond the scope of
this paper, and we defer a detailed study of the physics related to the
population of GW events to future work.}

% Rate analysis is now just counting, like what's done in phenomenological
% model, instead of forward modelling.
Working event-by-event as we do permits post-processing application of
arbitrarily complicated models of star formation and metalliticy evolution by
re-weighting the samples we obtain \citep[e.g.][]{VanSon2022}. Once we have
pulled the GW event posterior back from the observable space to the evolutionary
parameter space, \textbf{and have accounted for selection effects}, we can apply
the same hierarchical inference methods currently used to infer the compact
binary population to our progenitor population.  In particular, we can use our
progenitor population to infer the formation rate of compact binary population
progenitor systems over cosmic time without needing to make artifical and
simplistic assumptions about the delay time distribution or the
metallicity-specific star formation rate \citep{Vitale2019,Ng2021,VanSon2022}.
\textbf{In other word, just as we measure the event rate of GW events by
counting, we can measure the event rate of the progenitor systems of GW events
in the same way.} Comparing the formation rate of progenitors to the star
formation rate provides an avenue to check our understanding of binary evolution
and the relative rate of other formation channels.

>>>>

\begin{abstract}
One promising way to extract information about stellar astrophysics from
gravitational wave catalogs is to compare the catalog to the outputs of stellar
population synthesis modeling with varying physical assumptions.  The parameter
space of physical assumptions in population synthesis is high-dimensional and
the choice of parameters that best represents the evolution of a binary system
may depend in an as-yet-to-be-determined way on the system's properties.  Here
we propose a pipeline to simultaneously infer zero-age main sequence properties
and population synthesis parameter settings controlling modeled binary evolution
from individual gravitational wave observations of merging compact binaries. Our
pipeline can efficiently explore the high-dimensional space of population
synthesis settings and progenitor system properties for each system in a catalog
of gravitational wave observations.  We apply our pipeline to observations in
the third LIGO--Virgo Gravitational-Wave Transient Catalog. We showcase the
effectiveness of this pipeline with a detailed study of the progenitor
properties and population synthesis settings that produce mergers like the
observed GW150914.  Our pipeline permits a measurement of the variation of
population synthesis parameter settings with binary properties, if any;
we illustrate the possibility of such capability by presenting
inferences for the recent GWTC-3 transient catalog that suggest that the stable
mass transfer efficiency parameter may vary with primary black hole mass. 
\end{abstract}

@@

Traditionally, population synthesis studies simulate DCO populations with a
Monte-Carlo approach. Using theoretically motivated distributions of initial
parameters like age, metallicity, component mass, orbital separation, and
eccentricity, initial populations are evolved with fixed astrophysical
assumptions, or population synthesis hyperparameter settings, to produce a
synthetic catalog of DCO mergers which can be compared to parameterized models
derived from observations. This method has been applied extensively to
inference on GW data to inform phenomenological models
\citep[e.g.][]{Fishbach2017, Wysocki2019, Callister2021, Delfavero2021,
Farah2022, Delfavero2022}, population synthesis for a single formation
environment \citep[e.g.][]{Belczynski2016, Stevenson2017b, Taylor2018,
Barrett2018, Mastrogiovanni2022} or contributions from multiple formation
environments \citep[e.g][]{Zevin2017, Stevenson2017, Bouffanais2019, Zevin2021,
Wong2021, Bouffanais2021, Mapelli2022}. These approaches rely on several
assumptions, both in the simulations and model parameterizations fitted to the
GW data, which are unlikely to be fully correct. For example, without
knowing whether the parameterization of hyperparameters is sufficiently
accurate, there is no reason to suppose that the entire population of DCOs
should evolve with the same fixed hyperparameters; there is no a priori reason
the fitted parameterized model should capture the relevant features of the
observed DCO distributions to correspond to the binary physics of interest; etc.
A better approach would be to compare progenitor parameters and hyperparameter
settings directly to DCO observations, treating the population synthesis model
as a mapping from progenitor to merger parameters; merger parameters can then be
mapped into observations using a gravitational waveform family, and parameter
inference proceeds in the usual way, propagating information back up the mapping
chain \citep[e.g.][]{Veitch2015}.  A similar methodology, though with
\emph{fixed} hyperparameter settings, was advanced in
\citet{Andrews2018,Andrews2021}, as we discuss in more detail below.

@@

We find that the progenitor properties and even the ability to produce a
GW150914-like merger event is strongly correlated with hyperparameter settings,
consistent with results from prior forward modeling studies cited above;
different assumptions about the black hole masses at merger in GW150914 imply
wildly different formation channels and ZAMS progenitor masses for this event
and some nearby combinations of ZAMS masses are unable, with any reasonable
hyperparameter settings, to produce a DCO merger like GW150914.  We further
exhibit preliminary results of an ongoing analysis over the entire GWTC-3
catalog \citep{GWTC-3} that suggest that the accretion efficiency in stable mass
transfer may depend non-trivially on the primary black hole mass in merging
systems; our methodology allows for a systematic study of such dependencies,
which we will explore more fully in future work.

@@

\subsection{Statistical method}
\label{subsec:stats}
We define the progenitor's zero age main sequence (ZAMS) properties like mass,
orbital period, and eccentricity as progenitor parameters $\bm{\theta'}$, the
parameters that control different uncertain physical prescriptions such as wind
strength and common envelope efficiency as hyperparameters $\bm{\lambda}$, and
random variables that affect certain stochastic process such as whether CO natal
kick unbinds a binary as $\bm{X}$. To avoid clutter in the following derivation,
we denote all the parameters related to mapping a particular progenitor system
into a GW event collectively as evolutionary parameters $\bm{\Theta}$, which
includes $\bm{\theta'}$, $\bm{\lambda}$, and $\bm{X}$.

@@

\begin{align}
    p(\bm{\Theta} | d) = \frac{p(\bm{\theta}(\bm{\Theta})| d) \pi(\bm{\Theta})}{\pi_\mathrm{GW}(\bm{\theta}(\bm{\Theta}))}.
\end{align}
We apply a kernel density estimator with a Gaussian kernel \footnote{We use the
implementation in \texttt{scipy} \cite{2020SciPy-NMeth}} to the posterior
samples released in GWTC-3 to estimate $p(\bm{\theta}(\bm{\Theta})|d)$.
A desired prior in the evolutionary parameters can be included through
the prior $\pi(\bm{\Theta})$ in the above equation.

Including hyperparameters and random variables %in the sampling
drastically increases the dimensionality of the problem, which can make the
sampling process much more computationally expensive to converge. To speed up
the convergence, we first solve a root-finding problem to find points
$\bm{\Theta}$ that will generate GW parameters $\bm{\theta}$ close to the bulk
of the posterior density of a GW event.  We then use those points to initialize
a set of chains in the MCMC process. We use \textsc{emcee} \cite{emcee}
to sample the posterior density in the evolutionary parameter space.

@@

In principle, we should only accept solutions that exactly reproduce the LVK
posterior samples in the GW-observable space. However, it is not feasible to
achieve such a condition in practice, therefore we relax the condition in
eq.~\ref{eq:loss} to a small acceptance threshold. For this study, we picked a
threshold of $10^{-2}$ solar mass. For a more general study that use
observables that are not with the same unit, the norm should be chosen in a way
that all parameters are represented on a comparable scale. To make sure we find
a reasonably complete set of progenitor parameters that corresponds to the
posterior sample point, we use 1000 different initial guesses in solving the
optimization problem. As long as the solution fulfills the acceptance criteria,
the root is as valid as all other roots that fulfill the same acceptance
criteria, \emph{regardless to its initial guess}. This means, we have the
freedom to choose the set of initial guesses however we think might benefit the
optimization.

@@

\subsection{Binary evolution assumptions}
\label{subsec:binaries}

@@

We follow the critical mass ratio prescription 
described in \cite{Neijssel2019} for donors which fill their Roche lobes during
 other phases of stellar evolution.
For systems which begin a common envelope 
with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
stars merge since the donor has likely not formed a strong core-envelope boundary 
\citep{Ivanova2004, Belczynski2008}. We fix other assumptions for how stellar and binary 
evolution proceed. In particular, we assume that stars lose mass through stellar winds 
in a metallicity-dependent 
fashion according to \cite{Vink2001,Vink2005}. We assume that binaries are rapidly circularized
during both stable Roche-overflow mass transfer and common envelopes.
We further assume that the CO formation mechanism follows the delayed 
prescription of \citet{Fryer2012} which allows neutrino-driven supernova explosions to 
occur from instabilities which grow on timescales longer than $250\,\rm{ms}$, 
and that natal kicks due to supernovae are reduced by the amount 
of mass which falls back onto the proto-CO
as described in \citet{Fryer2012}.

\subsection{Case-specific configuration}
\label{subsec:configuration}

We apply the method described in section \ref{subsec:stats} to the
first GW event GW150914 \citep{GW150914}. We use the \texttt{Overall\_posterior}
posterior samples for GW150914, publicly available from the Gravitational wave
Open Science Center \citep{LIGOScientific:2019lzm}. The parameters involved in
the analysis and the range we allow for the inference are tabulated in table
\ref{tab:parameters}. The spin of BHs formed in isolated binaries depends
strongly on the angular momentum transport in massive stars which is still
highly uncertain (see however \citet{Fuller2019, Bavera2020}), therefore we only
consider the two component masses in the observable space $\bm{\theta}$ in both
the root-finding stage and the MCMC stage. Instead of the sampling the posterior
in the two component masses space, we choose to perform the KDE in the chirp
mass and mass ratio space, which are less degenerate and is less prone to error
from the KDE.

For progenitor parameters $\bm{\theta'}$, we characterize each guess with five
parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the
orbital period $t_{\rm orb}$, the eccentricity $e$, and the metallicity $Z$ at
ZAMS. We use a log-uniform prior in sampling $M_{\rm 1,ZAMS}$, $M_{\rm
2,ZAMS}$, $t_{\rm orb}$, and $Z$, and we use a uniform prior for $e$.
The range follows the range for the optimization range as given by table
\ref{tab:parameters}. Note that the progenitor formation redshift is not
obtained through the root-finding process because the formation redshift does
not intrinsically affect the evolution of the binary. Instead, we compute the
time it takes for the binary to evolve from formation to the merger and add this
time to the lookback time when a GW event is observed and limit the total time
to merger to be less than $13.7\,\rm{Gyr}$. Using the total lookback time, we
can compute the redshift of the binary at ZAMS formation following the Planck
2018 cosmological model \citep{Planck2018}. The main benefit of post-processing
the redshift in this way is that we do not rely on a particular assumption of
star formation rate (SFR) distribution when we solve for the posterior
distribution in redshift. This means we can put the prior on formation redshift
(or in general, metallicity-redshfit distribution) in by reweighting the
posterior distribution according to a particular SFR model, and therefore do not
have to rerun the inference every time we change the SFR model. This is
especially helpful since SFR models are still highly uncertain but strongly
impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.

For hyperparameters $\bm{\lambda}$, we choose parameters that strongly affect
binary evolution in \texttt{COSMIC} for massive stars, including the common
envelope efficiency $\alpha$, the fraction of mass accreted onto the accretor
during stable mass transfer $f_{\rm acc}$, the critical mass ratio which
determines whether Roche-overflow mass transfer proceeds stably or produces a
common envelope when the donor star is on the Hertzsprung gap $q_{\rm crit,3}$,
and the root-mean-square of the Maxwellian distribution employed for CO natal
kicks, $\sigma$. We use a uniform prior for $\alpha$, $f_{\rm acc}$, and
$q_{\rm crit,3}$, and a log-uniform prior for $\sigma$.

We use the same prior for all the events in this study. Note that our
main goal is to see whether the events are compatible with \textsc{COSMIC} on an
individual event basis, we do not consider any population analysis here, and
therefore we do not account for any selection effect.

@@

We show a portion of the joint posterior of the progenitor parameters and
hyperparameters of GW150914 in figure \ref{fig:GW150914_posterior}. Each panel
is colored based on the comparison between progenitor parameters (blue),
hyperparameters (orange), or a mix of the two (green). The choice
for $\sigma$ does not affect the formation of GW150914-like BBHs. This is not
unexpected since the \citet{Fryer2012} delayed model can significantly reduce the
natal kick strength at BH formation based on the amount of fallback onto the
proto-CO. In the case of GW150914-like BBHs, the BH progenitors are massive
enough that the fallback reduces the natal kick to zero. Because of this we do
not include $\sigma$ in Figure~\ref{fig:GW150914_posterior}. Similarly, we do
not include the ZAMS orbital period and eccentricity, which are correlated with
one another, but not strongly with the other progenitor parameters or
hyperparameters.

The formation redshift distribution is slightly correlated with the $q_{\rm
crit, 3}$ and $f_{acc}$ hyperparameters, though the GW150914-like progenitors
generally prefer formation redshifts of $z\sim0.1$ (delay times of
$\lesssim0.1\,\rm{Gyr}$). This correlation is due to the preference for longer
delay times between ZAMS formation and merger for progenitors which undergo
stable mass transfer at Roche overflow instead of a common envelope. This result
is in direct agreement with \citet{vanSon2022a} who also find that the common
envelope evolution channel produces BBH mergers with shorter delay times than
the stable Roche-overflow channel. \textbf{This result is also consistent with
literature that doesn't consider the delay times originating from common
envelope and stable mass transfer separately, but is dominated by
short-delay-time mergers originating from common envelope
\citep[e.g.][]{OShaughnessy2010, Dominik2012, Eldridge2016, Mapelli2017,
Lamberts2018, Broekgaarden2022}.}

The joint posterior exhibits strong correlations between the ZAMS masses, as well as
strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$, and
$f_{\rm{acc}}$ and $\alpha$. We show scatter plots of each of these combinations
in the upper-right inset of the figure.

The correlations between the ZAMS masses are similar to those found by A21 with
the majority of the population preferring primary and secondary masses between
$60$--$90\,M_{\odot}$. However, we also find some ZAMS masses which extend up to
$150\,M_{\odot}$ which is the limit imposed by our assumptions. The features in the scatter points of the ZAMS masses, most obvious in the region where $M_1,\rm{ZAMS}\sim 140-150\,M_\odot$, are the outcome of the interplay between the progenitor parameters and hyperparameters. The dominant cause of regions with fewer data points is that fewer binary progenitors produce BBH mergers with chirp masses and mass ratios consistent with GW150914's priors. This is most clearly illustrated for progenitors with primary masses between $142-144\,M_\odot$ and secondary masses below $80\,M_[\odot]$. We explore the
formation scenarios which lead to successful GW150914-like mergers and those
which fail to produce GW150914-like mergers in more detail below in
Figure~\ref{fig:compare_fixed_variable}.

@@

% Paragraph discussing the potential application on the population side.
To illustrate the potential benefit of using our method on the population level,
we perform the same analysis for all events in GWTC-3. For events announced in
GWTC-1, we use \texttt{Overall\_posterior},
\texttt{PublicationSamples},\texttt{IMRPhenomXPHM\_comoving} and
\texttt{C01:Mixed} posterior samples, respectively. All the data are publicly
available from the Gravitational wave Open Science Center
\citep{LIGOScientific:2019lzm}. Plotted in figure \ref{fig:GWTC-3_f_acc_mass} is
the posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
most\footnote{We excluded events for which our reconstructed posterior has KL
divergence greater than 0.1 nats relative to the GWTC-3 posterior; see below.}
of the events in GWTC-3. Each contour represents the $68\%$ credible interval of
the posterior density for that particular event. There is a suggestive trend
showing that $f_{\rm acc}$ could increase as the mass of the progenitor
increases. Such a trend implies the stable mass transfer phase of a less massive
binary would be preferentially non-conservative, with more conservative mass
transfer in more massive binary systems.

We have validated the robustness of every events shown in
\ref{fig:GWTC-3_f_acc_mass} by reprojecting the posterior density in the
evolutionary parameter space to the observable space. Note that some events did
not pass the KL divergence test we proposed in section \ref{sec:method}; we do
not include these systems in figure \ref{fig:GWTC-3_f_acc_mass}. Binaries that
form low-mass compact objects have lower amounts of fallback and tend to have
correspondingly larger variance in their random natal kicks. Larger kicks can
unbind the progenitor binary, or lead to wide binaries that do not merge within
a Hubble time.  In these cases, the extra variance during the reprojection can
produce a posterior that may not agree with the original posterior, hence
yielding a higher KL divergence.

@@

We present a new pathway to understand binary evolution with GW events in this
paper. Instead of forward modelling an assumed distribution of initial binaries
to the observed population, we find the corresponding evolutionary parameters
event-by-event. In this first work, we showcase the power of the proposed method
with an application to GW150914. We show the joint posterior of the event's
progenitor parameters and hyperparameters.  Our work is a data-driven
way to study GW event progenitors, and also allows the possibility of
constraining the astrophysics related to binary evolution, especially by
capturing the correlation between hyperparameters in different systems.

Our results suggest that the accretion efficiency during stable mass transfer
may depend on the primary black hole mass.  In general, our method returns the
joint posterior distribution of progenitor parameters and hyperparameters for
each event, which enables a data-driven way to study the distribution of
hyperparameters.  That is, once we have a catalog of events, each has their own
posterior in the hyperparameter space, and we can employ well-known techniques
such as hierarchical Bayesian analysis to fit a population model to the
distribution of hyperparameters; this avoids making overly-specific assumptions
like fixing the hyperparameters for all types of event. In this work we
focus on illustrating the possibililty of conducting such analysis with our
pipeline. Note that there are concerns that need to be addressed before one can
make statments about the physics, such as the number of channels considered and
specific choices of parameterization. These concerns are beyond the scope of
this paper, and we defer a detailed study of the physics related to the
population of GW events to future work.

% Rate analysis is now just counting, like what's done in phenomenological
% model, instead of forward modelling.
Working event-by-event as we do permits post-processing application of
arbitrarily complicated models of star formation and metalliticy evolution by
re-weighting the samples we obtain \citep[e.g.][]{VanSon2022}. Once we have
pulled the GW event posterior back from the observable space to the evolutionary
parameter space, and have accounted for selection effects, we can apply
the same hierarchical inference methods currently used to infer the compact
binary population to our progenitor population.  In particular, we can use our
progenitor population to infer the formation rate of compact binary population
progenitor systems over cosmic time without needing to make artifical and
simplistic assumptions about the delay time distribution or the
metallicity-specific star formation rate \citep{Vitale2019,Ng2021,VanSon2022}.
In other words, just as we measure the event rate of GW events by
counting, we can measure the event rate of the progenitor systems of GW events
in the same way. Comparing the formation rate of progenitors to the star
formation rate provides an avenue to check our understanding of binary evolution
and the relative rate of other formation channels.
